---
title: "creating an R data package"
author: "Davinder Singh"
date: "2023-04-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## creating a new repository locally as an R package

With Rstudio
1. Use option to create a new project
2. Choose create an R package
3.Choose create an R package
4. Give a name to the package
5. Click on create in new session

Once we have a backbone for an R package, start customizing it 
- Modifying the descrption file
- This files contains the metadata of the package
Metadata is "data about data", it contains the information
licenese GPL (>=2) is the most common for data package
Say something about the particularities 

### Add a readME file

Use a new package that is called 'usethis'
```{r eval=FALSE}
#install.packages("usethis")
library(usethis)
```
```{r}
use_readme_md()
```
### Adding data to the package

We are going to use the function 'use_data_raw()', this creates a folder called 'data-raw/' with an R script tht will document the creation of the data object that will be stored in our package.

In this R script called 'DATASET.R' we will document the steps to read the raw data which is in CSV format and save it in R format
This is useful so that the dataset are always available when someone installs an R package.

### Documenting the data (and functions)

Help for a function or data set can be accessed with the question mark '?', or the function 'help()'
But where are these documentation files created?

There is a special folder called 'man/' in every R package that stores files of type '.Rd', shortcut for 'Rdocument.'
To render from markdown in R script to .Rd in the man folder, we use the function 'document()' from the package devtools:
```{r eval=FALSE}
#install.packages("devtools")
library(devtools)
document() #run this line in the repo of your R package
```
## Building a website for an R package

We will use the 'pkgdown' package.
We first activate the use of pkgdown in our repository with use 'use_pkgdown()' from package 'usethis'
```{r}
library(pkgdown)
build_site()
```

## Customize the website

## Adding vignettes to our package

Vignettes are Rmd files that have information on how to use the package.
To create a new one from scatch we use the function 'use_vignette()' from the package 'usethis'
```{r eval=FALSE}
library(usethis)
use_vignette(name = "intro")
```

