---
title: "functions-prep"
author: "Davinder Singh"
date: "2023-03-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Writing your own function in R

## writing a function how it works
```{r}
function_name <- function(inputs) {
  ouput_value <- do_something(inputs)
  return(output_value)
}
```

## example of brackets
```{r}
{
  a =2
  b =3
  a + b
}
```
## example of how function work to find volume
```{r}
calc_shrub_vol <- function(length, width, height) {
  area <- length * width
  volume <- area * height
  return(volume)
}
calc_shrub_vol(0.8, 1.6, 2.0)
shrub_vol <- calc_shrub_vol(0.8, 1.6, 2.0)
```

# How Functions Excute

## function creates its own internal environment once code is complete you can't use anything from the function
```{r}
calc_shrub_vol <- function(length, width, height) {
  area <- length * width
  volume <- area * height
  return(volume)
}
calc_shrub_vol(0.8, 1.6, 2.0)
shrub_vol <- calc_shrub_vol(0.8, 1.6, 2.0)
```

# Setting Default Values for Arguments

## if you want to set a value for a function set it a value. ex (a,b, c =  )
```{r}
calc_shrub_vol <- function(length, width, height =1) {
  area <- length * width
  volume <- area * height
  return(volume)
} # in video they did height = 1 to set the height to 1 meter
calc_shrub_vol(0.8, 1.6, 2.0)
calc_shrub_vol(0.8,1.6) # since height is not added the default function is multuplied in this case 1
```

# When to use named and unnamed arguments

## when assigned it gets same value because of function
```{r}
calc_shrub_vol <- function(length, width, height =1) {
  area <- length * width
  volume <- area * height
  return(volume)
}
calc_shrub_vol(0.8, 1.6, 2.0)
calc_shrub_vol(0.8,1.6)
calc_shrub_vol(length = 0.8, width = 1.6, height =  2.0)
calc_shrub_vol(height =  2.0, length = 0.8, width = 1.6) # cause function is l * W * H
calc_shrub_vol(0.8, 1.6, height = 2.0) # change height
```

# Combining Functions

## another way to get mass using intermediate variables
```{r}
est_shrub_mass <- function(volume) {
  mass <- 2.65 * volume^0.9
  return(mass)
}
shrub_volume <- calc_shrub_vol(0.8, 1.6, 2.0)
shrub_mass <- est_shrub_mass(shrub_volume)
```

## pipes another way 
```{r}
shrub_mass <- calc_shrub_vol(0.8, 1.6, 2.0) %>% 
  est_shrub_mass()
```

## nesting functions
```{r}
shrub_mass <- est_shrub_mass(calc_shrub_vol(0.8, 1.6, 2.0))
```

# Calling functions from inside other functions

## using functions inside other functions example
```{r}
est_shrub_mass_dim <- function(length, width, height = 1) {
  volume <- calc_shrub_vol(length, width, height)
  mass <- est_shrub_mass(volume)
  return(mass)
}
est_shrub_mass_dim(0.8, 1.6, height = 2.0)
```


# Rstudio tips and tricks for working with functions

## Top level to jump around code

## arrows to collapse code to hide code

## highlight function in tools -> global options -> display "Highlight R function calls" functions calls are gray, variables are black, key words are blue
