---
title: "if-else conditions"
author: "Davinder Singh"
date: "2023-03-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## logical and conditional statements

These are pieces of code that return the 'TRUE' or 'FALSE' values, that is, a logical value

The common operators of lofical statementsa are:
-equality '=='
-inequality '!='
-greater than '>'
-less than '<'
- greater or equal than '>='
-less or equal than '<='

The conditonal statements allow to test several logical conditions at a time
The condition operators (or symbols) are
-AND '&' (inside dplyr functin we can represent AND using a ',')
-OR '|'

We also have logical functions that test if something is 'TRUE' or 'FALSE', for example:
-'is.na()' is a function that tests if a value is an 'NA'
-This function i part of a whole family of functions, they also start with 'is.':
-is.vector()
-is.data.frame()
-is.factor()

For next class: how to get all functions from a family (method).

-which() : takes logical vectors, it will give you the numerical index (position)
of all values that are TRUE
```{r}
which(letters == "r")
letters[18]
```
## excerise 6
```{r}
 w <- 10.2
 x <- 1.3
 y <- 2.8
 z <- 17.5
 colors <- c("red", "blue", "green")
 masses <- c(45.2, 36.1, 27.8, 81.6, 42.4)
 dna1 <- "attattaggaccaca"
 dna2 <- "attattaggaacaca"

```

```{r}
w > 10
colors == "green"
x > y
masses > 40
2^x + 0.2 == y
dna1 == dna2
dna1 != dna2
w > x | y > z
x * w < 13.5
x * w > 13.2
x * w < 13.5 & x * w > 13.2
masses > 30 &  50
masses > 30
masses < 50
masses < 50 & masses > 30
```

## How to make simple choices with 'if()'
```
if(condition is TRUE) {
  Run all lines of code in this block of code
}
```
If the condition is not TRUE, then nothing happens.

## Excerise 7: Handling one choice
```{r}
 age_class = "sapling"
 if (age_class == "sapling"){
y <- 10
 }
 y
```

## Case when we have two options: if-else structure

The general form of this structure:

```
if(condition){
code that runs if condition is met
} else{
code that runs if condition is NOT met
}
```
## Excerise 8 handling two choices
```{r}
 age_class = "seedling"
 if (age_class == "sapling"){
y <- 10
 } else{
   age_class == "seedling"
   y <- 5
 }
 y
```

## handle more than 2 choices

In this case we are using the elseif structure:
```
if(condition1){
first block code that is execities if conditon 1 is met
} else if(condition2){
second block code taht excutes if condition2 is met
} else if(conditon3){
more code
} else{
this will cover all the conditoins that are not specified before
}
```

you do not have to end up with and else block
'Else if' are more intentional with the conditions
A simple else will run in all other no matter what.

## Exceris 9 handling 3 choices or more 
```{r}
 age_class = "adult"
 if (age_class == "sapling"){
y <- 10
 } else if(age_class == "seedling"){
   y <- 5
 } else{
   y <- 0
 }
 y
```

## Excerise 12 load or download file
```{r}
list.files()
if (file.exists("surveys (1).csv")) {
  print("file exists")
} else{
  print("file does not exist")
}
```
?download.file
```{r}
if (file.exists("surveys.csv")) {
  print("file exists")
} else{
  download.file("https://ndownloader.figshare.com/files/2292172", "surveys_test.csv")
  read.csv("surveys_test.csv") %>% 
  head() %>% 
    print()
  
}
```
```{r}
if (file.exists("species1.csv")) {
  print("file exists")
} else{
  download.file("https://ndownloader.figshare.com/files/3299483", "species.csv")
  read.csv("species.csv") %>% 
  head() %>% 
    print()
  
}
```



