---
title: "Joining-Vectors"
author: "Davinder Singh"
date: "2023-03-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## What are vectors?

they are unidemnsional matrices
they can only hold one type of data, either numeric (integer or double),character, or logical(complex numbers)

## What are data frames

It is a two dimentisnoal matix: rows, and columns
it can hold any type of data
it can only hold different types of data in a certain way:
only column can have different data types
within a column, all rows must have same data type

A data frame can also be defined as a collection of vectors (they can be of different or the same type) all of same length!

```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
str(surveys)
```

## creating vector
```{r}
c("luna", "avi", "Anita", "James", "Charles", "Damian", "Davinder") -> our_names
str(our_names)
1:7 # the colon operator creates a vector of numbers
1:7 -> my_numbers
-100:200
```
In how many steps or units does the colon operator increase? It increases in a step of 1.

What do we do if I won to create a numeric sequence that increases in steps different than 1?
?seq
```{r}
seq(-100, 200, by = 0.1) -> my_numbers
str(my_numbers)
```
```{r}
letters
LETTERS
```


## Creating data frames from vectors

the most general way to do this is with the function 'data.frame()':

```{r error=TRUE}
data.frame(names = our_names, numbers = my_numbers)
```
Rember: vectors must have the same length to be part of a data.frame!
must be a multiple of vector

```{r}
data.frame(names = our_names, numbers = 1:7)
```

Even though we can't use vectors that are larger to create a data frame, we can use vector that or eshorter, and R will recycle the values

```{r}
data.frame(names = our_names, numbers =1)
```
R will recycle the value only if they are multiple of vector of names two times, so it is a multiple of 2.
```{r}
data.frame(names = rep(our_names, 2), numbers = c(2, 5.5))
```
?data.frame
## Excerise 6
```{r}
length <- c(2.2, 2.1, 2.7, 3.0, 3.1, 2.5, 1.9, 1.1, 3.5, 2.9)
width <- c(1.3, 2.2, 1.5, 4.5, 3.1, NA, 1.8, 0.5, 2.0, 2.7)
height <- c(9.6, 7.6, 2.2, 1.5, 4.0, 3.0, 4.5, 2.3, 7.5, 3.2)
data.frame(length, width, height, genus = "Taxus", species = "epithet baccata")
```

## Extracting/accessing values from vectors and data frames

```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
str(surveys)
```
One common way to extract or access vectors crom column in a data frame is the '$'
```{r}
surveys$record_id -> record_id
```

Another way is with the square brackets '[]'

```{r}
surveys[1:10, "hindfoot_length"]
```

If I want all the values from the rows of column hindfoot length:

```{r}
surveys[, "hindfoot_length"] -> hindfoot_length
```

Another way is using double square brackets
```{r}
surveys[["record_id"]] %>% 
  head()
```

```{r}
surveys$weight -> surveys_weight
surveys[, "month"] -> surveys_month
surveys[, "hindfoot_length"] 
mean(hindfoot_length, na.rm = TRUE)
```

