---
title: "joining-tables"
author: "Davinder Singh"
date: "2023-03-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

Load the three data sets that we are going to join, survey.csv, speices.csv, plot.csv
```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
species <- read.csv(file = "../data-raw/species.csv" )
plots <- read.csv(file = "../data-raw/plots.csv")
```

## Why do we need to combine or join dat tables

homework: elaborate on this topic

## How do we join data tables in R

There is a group function '-join()' that allow us to combine two data tables using values on a shared column

there has to be a shared column, and we need three main arguments to run these functions, two data tables and one column name

The different function allow us to combine in a differnt ways.
'inner_join'

```{r results='hide'}
inner_join(surveys, species, by = "species_id")
```

We can also run it using pipes:
```{r results='hide'}
surveys %>% 
  inner_join(species, by = "species_id") -> joined_table
```

### How can we explore our combined/joined head table?

We want to see differences between the two input tables to see difference in columns we can use 'head()'
To see number of rows we can use 'str()'
```{r}
head(species)
head(surveys)
head(joined_table)
str(species)
str(surveys)
str(joined_table)
```

what happened with the number of rows in joined_table vs surveys?

It dropped the rows that did not have matching vlaues of species_id column

## Excerise 1

```{r}
plots %>% 
  inner_join(surveys, by = "plot_id") %>% 
  filter(plot_type == "Control") %>% 
  head()
```

## Automate joining tables and other things with 'intersect()' 

Which species_id values are shared between the two data tabels
```{r}
intersect(surveys$species_id, species$species_id)
```

To find shared columnss we use 'colnames()' function
## Excerise 2
```{r}
colnames(surveys)
colnames(species)
intersect(colnames(surveys), colnames(species))
```
```{r }
colnames(plots)
colnames(surveys)
intersect(colnames(plots), colnames(surveys))
plots %>% 
  inner_join(surveys, by = "plot_id") %>% 
  filter(plot_type == "Rodent Exclosure") %>% 
  head()
```

## other join functions

'left_join()' retains all values from the first table, drops unmatching from second

'right_join()' drops values from the first table and retaining all values from second

'full_join()' keeps all values from both tables

## Joining multiple data tables
can we '_join()' function on 3 or more table at same time?
NO
so we use a pipe on call the join function two or more times (as needed):
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  str()
  
```
## Excerise 3
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  filter(plot_type == "Long-term Krat Exclosure" | plot_type == "Control") %>% 
  filter(taxa == "Rodent") %>% 
  filter(!is.na(weight)) %>% 
  select(year, genus, species, weight, plot_type) %>% 
  str()
  
```

## Excerise 4
## help on 3, 5 and 6 and ex 5
### Ex 4 p.1
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  select(year, month, day, species_id, weight) %>% 
  filter(species_id == "DO") %>% 
  head()
```
### problem 2
Create a data frame with only data for species IDs "PP" and "PB" and for years starting in 1995, with the columns "year", "species_id", and "hindfoot_length", with no missing values for "hindfoot_length"
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  select(year,species_id, hindfoot_length) %>% 
  filter(year <= "1995", !is.na(hindfoot_length)) %>% 
  filter(species_id == "PP" | species_id == "PB") %>% 
  head()

```
### problem 3
Create a data frame with the average "hindfoot_length" for each "species_id" in each "year" with no null values.
```{r}
surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  group_by(year, species_id) %>% 
  summarize( hindfoot_length = mean(hindfoot_length, na.rm = TRUE)) %>% 
  head()
```

### excerise 4
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  select(year, genus, species, weight, plot_type) %>%
  filter(genus == "Dipodomys") %>% 
  head()
  
```

```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  ggplot(mapping = aes(x = weight, y = hindfoot_length)) +
  geom_point(mapping = aes(color = species_id)) +
  scale_x_log10()
```
### excerise 6
Make a histogram of weights with a separate subplot for each "species_id". Do not include species with no weights. Set the "scales" argument to "free_y" so that the y-axes can vary. Include good axis labels.
?geom_histogram
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  filter(!is.na(weight)) %>% 
  ggplot() + 
  geom_histogram(mapping = aes(x = weight, color = species_id)) +
  facet_wrap(~species_id, scales = "free_y")
```

## Excerise 5
?order
The table should be sorted first by the species (so that each species is grouped together) and then by weight, with the largest weights at the top.
```{r}
inner_join(surveys, species, by = "species_id") %>% 
  inner_join(plots, by = "plot_id") %>% 
  select(month, day, year, species_id, weight, hindfoot_length) %>% 
  filter(!is.na(weight)) %>% 
  arrange(species_id, desc(weight)) %>% 
  head()
```

