---
title: "Tree opentree"
author: "Davinder Singh"
date: "2023-04-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ape) #Analysis, Phylogentics and Evolution R package
#install.packages("rotl")
library(rotl) # an R interferface fpr open Tree of life
library(remotes)
library(datelife)
```

Insect - 1 million named species
Mammals - 5k named species 
Flowering plants - 200k named species
### Read a newick tree from open tree of life

```{r}
tree <- read.tree(file = "../data-raw/subtree-node-mrcaott9245ott242638-Salamandroidea--Sirenoidea (1).tre")
```
```{r, fig.height= 20, fig.width= 15}
plot(tree, cex = 0.2)
mtext(text = "amphibian-environmental")
```

We can also use the url of a tree to read a tree into R. We will provide the URL in the argument 'file ='
Tje is "http://ape-package.ird.fr/APER/APER2/primfive.tre":
```{r}
small_tree <- read.tree(file = "http://ape-package.ird.fr/APER/APER2/primfive.tre")
small_tree
```

```{r}
plot(small_tree, cex = 0.5)
```
The R structure of a phylogentic tree

###Getting a tree for your species from the open tree of life

The Open tree of life has tools that allow to match the name of a group or species to their unified taxononmy so that we can look for information

If we are doing this in R, we are using the 'tnrs_match_names()' from the package 'rotl':
```{r}
tnrs <- tnrs_match_names(names = "Homo")
class(tnrs)
```

What is the main difference between the information that the information that we have in the column 'search_string' and 'unique_name':
One is the search string and the other is the scientific taxonomci make from the open tree of life unified taxononmy (OTT)

We can use the unique number identifier from OTT to get information from that taxon.

To get a tree we can use the function 'tol_subtree()'
```{r}
#tol_subtree(ott_id = tnrs$ott_id)
```

### Getting a tree for a set of taxa

Sometimes we have a very large groups and we just want to get a small tree representing some lineages insdie the group

For example, we want a tree of all orders in amphibians. we known the common names of these which are frogs, salamnders, cecilinas

We want to find this lineages in the Open tree of life; the steps to do this are:
1. Get the scientific names of our taxa
```{r}
c("anura","urodela","gymnophiona") -> amphibians_orders
```

```{r}
tnrs_match_names(names = amphibians_orders) ->amph_order_tnrs
amph_order_tnrs
```
We have two colums at the beginning, search string contains the names that we provided, and unique_name contains the most updated or offical or most commonly used taxonomic names for that we provided.

3.Next step is to use the OTT ids that we got to extract the tree containing those lineages.
If we have only one name or lineage, we can use 'tol_subtree()'
If we have more than one names, we will use 'tol_induced_subtree()'

```{r}
tol_induced_subtree(ott_ids = amph_order_tnrs$ott_id) -> amp_order_tree
amp_order_tree
```
```{r}
plot.phylo(x = amp_order_tree)
```

### Getting a tree of all families

Install an'datalife' R package from Github, with a function


```{r}
install_github(repo = "phylotastic/datelife", force = TRUE)

```
```{r}
#install.packages("remotes")
install.packages("datelife")
```

Now we can use the 'datalife' package, but we have to load it first
```{r}
amph <-tnrs_match_names(names = "Amphibians")
tol_subtree(ott_id = amph$ott_id)
get_ott_children(ott_ids = amph$ott_id, ott_rank = "family") -> amphibian_families
amphibian_families
```
```{r}
tol_induced_subtree(ott_ids = amphibian_families$Amphibia$ott_id)
```


we are going to use the fucntion that extracts all the names from a taxon that belong to a given named rank. The function is called 'get_ott_childern()':

HW:
1. extract the names of the taxa from your data set for the final project
2. Match them OTT using TNRS with 'tnrs_match_names()'
3. Use the OTT ids to get an induced subtree
