---
title: "Tree join and viz"
author: "Davinder Singh"
date: "2023-04-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(BiocManager)
library(ggtree)
library(ape)
library(dplyr)
```

## The package 'ggtree'

There are a couple of databases for hosting biological software in R, the most famous one is CRAN, but another very reliable one is BioConductor

When we are installing package from CRAN, we use the function install.package(). To install packages from Bioconductor we use a function from an R package that is called 'BiocManger'

install.packages("BiocManager")

Once that this pacakge is installed, we can use one of its functions that is called 'install()' to install packages from Bioconductor directly: 'install("ggtree")'
```{r}
portal_tree <- read.tree(file = "../data-raw/portal-tree.tre")
```

```{r}
ggtree(portal_tree)
```
```{r fig.height= 10, fig.width = 8}
plot.phylo(portal_tree, cex = 0.5)
```
### The structure of a tree in R 

We have an object that we named 'portal_tree', how do we explore its structure? The 'str()' function is a great option
What is the calss and length of this object
```{r}
str(portal_tree)
```

```{r}
class(portal_tree)
```

How would you define a 'phylo'
```{r}
taxonomy <- read.csv(file = "../data-raw/portal-species-taxonomy.csv")
head(taxonomy)
```
```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
species <- read.csv(file = "../data-raw/species.csv")
intersect(colnames(surveys), colnames(species))
joined <- inner_join(surveys, species, by = "species_id")

```

```{r}
left_join(portal_tree)
```


