---
title: "visualization-uhuru-day2"
author: "Davinder Singh"
date: "2023-02-28"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval = TRUE}
getwd()

```

```{r eval = TRUE}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

``` {r eval = TRUE}
acacia <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```

```{r}
acacia$HEIGHT <- as.numeric(acacia$HEIGHT)
acacia$HEIGHT
```

```{r}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt",
         sep = "\t",
         na.strings = "dead")
is.numeric(acacia$HEIGHT)
```

```{r}
plot(x = acacia$HEIGHT, y = acacia$AXIS1, main = "Acacia height vs axis1")
```

```{r}
library(ggplot2)

ggplot(data = acacia, mapping = aes(x = HEIGHT, y = AXIS1, color = TREATMENT)) +
  geom_point(size = 5) +
  labs(x = "Tree height", y = "Tree axis 1", title = "Acacia trees size relationships")

```
## rescaling axes
```{r}
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT)) +
geom_point(size = 3, color = "blue", alpha = 0.5) +
scale_y_log10() +
scale_x_log10()
```

``` {r}
ggplot(acacia, aes(x = CIRC, y = HEIGHT)) +
geom_point(size = 3, alpha = 0.5) +
geom_smooth()
```
# what happens when we fit a linear model to data grouped by treatment?
```{r}
ggplot(acacia, aes(x = CIRC, y = HEIGHT, color = TREATMENT)) +
geom_point(size = 3, alpha = 0.5) +
geom_smooth()
```
What happens is that there is a line of best fit and the data for each treatment

### line type
```{r}
ggplot(acacia, aes(x = CIRC, y = HEIGHT, color = TREATMENT, linetype = TREATMENT)) +
geom_point(size = 3, alpha = 0.5) +
geom_smooth()
```
### method = geom_smooth()
```{r}
ggplot(acacia, aes(x = CIRC, y = HEIGHT, color = TREATMENT)) +
geom_point(size = 3, alpha = 0.5) +
geom_smooth(method = "lm") # try with "glm"
```

## geom_bar
```{r}
ggplot(data = acacia, mapping = aes(x = TREATMENT)) +
geom_bar()
```

## geom_histogram()
```{r}
ggplot(acacia, aes(x = CIRC)) +
geom_histogram(fill = "red")
```
## proportion of data colored by treatment
```{r}
ggplot(acacia, aes(x = CIRC, color = TREATMENT)) +
geom_histogram(bins = 20)
```
## assign color with gray outline
```{r}
ggplot(acacia, aes(x = CIRC, fill = TREATMENT)) +
geom_histogram(bins = 20, color = "gray")
```

## change these values across layers
```{r}
ggplot() +
geom_point(data = acacia,
           mapping = aes(x = CIRC, y = HEIGHT,
                         color = TREATMENT)) +
geom_smooth(data = acacia,
            mapping = aes(x = CIRC, y = HEIGHT))
```

## plot categorical and continous data on same plot
```{r}
ggplot() +
geom_point(data = acacia,
           mapping = aes(x = CIRC, y = HEIGHT,
                         color = TREATMENT)) +
geom_smooth(data = acacia,
            mapping = aes(x = CIRC, y = HEIGHT)) +
geom_histogram(data = acacia,
               mapping = aes(x = CIRC, color = TREATMENT),
               alpha = 0.1)
```

#gg save
```{r}
ggsave("acacia_by_treatment.jpg")
```

## size of image in inches
```{r}
ggsave("acacia_by_treatment.pdf", height = 5, width = 5)
```