---
title: "UHURU data set visualization"
author: "Davinder Singh"
date: "2023-02-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## A tip for working with Rmarkdown
The Working directory inside this Rmarkdown r chunk is the following:
``` {r eval = TRUE }
getwd()
```


Note: remember that workingdirectories in a project and an R chunk are not always the same!

# 1. Describing the working data set

UHURU data set
The UHURU(ungulate herbivory under rainfall uncertainty) aim to characterize the effects of successively removing the largest size classes of herbivores from the system as well as evaluate how the direction and magnitude of these effects are shaped  by variation in precipation regimes
## 2. Reading the data set
We are reading a data set in TSV format. This is a "tab separted values" file.
To read it into R we use the function `read.csv()` with the `sep =` argument set to `"\t"` which represents a tab in computer language.
``` {r eval = TRUE}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```
Assign the data to a varibale so we can work with it
``` {r eval = TRUE}
acacia <- read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt", sep = "\t")
```
### 3. Acessing elements of a `data.frame`
```{r}
head(acacia)
str(acacia)

```

It is similar to what we do for vectors, but we have two dimensions
```{r}
acacia[,6]
str(acacia)
```
All are the following ways to accesing a column or variable in data
```{r}
acacia$SURVEY
acacia[, 1]
acacia[, "SURVEY"]
```
### Replacing elements in a data frame
```{r}

```

## 4. quality control
```{r}
head(acacia)
str(acacia)
acacia
class(acacia$HEIGHT)
is.numeric(acacia$HEIGHT)
acacia$HEIGHT
as.numeric(acacia$HEIGHT)
```
We identified that HEIGHT should be numeric and is instead character:

```{r}
acacia$HEIGHT <- as.numeric(acacia$HEIGHT)
acacia$HEIGHT
```

Coercion in R computer language ,eams tjat a va;ie was forced to be a type, in this case, the word "DEAD" cannot be transformed to a number, so it is assigned the NA value

```{r}
read.csv(file = "../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt",
         sep = "\t",
         na.strings = "dead")
is.numeric(acacia$HEIGHT)
```

## 5. Plotting data with `ggplot2`

Regular way to plot in R

```{r}
plot(x = acacia$HEIGHT, y = acacia$AXIS1, main = "Acacia height vs axis1")
```

With ggplot, we create layers

```{r}
library(ggplot2)

ggplot(data = acacia, mapping = aes(x = HEIGHT, y = AXIS1, color = TREATMENT)) +
  geom_point(size = 5) +
  labs(x = "Tree height", y = "Tree axis 1", title = "Acacia trees size relationships")

```

```{r}
colors()
```

### experimental treatment and goal
The goal experimental treatment is to find which treatment is the best and growing the largest trees.

### different variables that were measured
the different variables that were measured were mega, meso, open, and the total.

### what is the influnce of experimental treatment?
The influence of the experimental treatment is the meso treatment

### what happened to the acacias in the open plot?
The open treatment didn't show much growth comprared to other treatment groups as meso, and mega had trees of larger height
