---
title: "Data manipulation with pipes"
author: "Davinder Singh"
date: "`2023-03-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(magrittr)
```

## The calssic way of running code

For example I want the square root of the mean of a sequenceo of numbers

### Nested code

```{r}
numbers <- 1:300
mean(numbers)
sqrt(mean(numbers))
```

### Sequential code

In this case we create intermediate varaibles
```{r}
numbers <- -300:546
numbers_mean <- mean(numbers)
sqrt(x = numbers_mean)
```

## Piping code

It can be implemented in R using the package 'magrittr'. It is a dependecny of 'dpylr', so it is installed along

```{r}
library(magrittr)
```

The symbol of the pipe is '%>%',
But we also have a new symbol that is similar to bash '|>'
The purpose of pipes is to elimante or reduce to the max the need of intermediate variabels

```{r}
1:300 %>% mean()%>% sqrt()
```

### Pipes with the surveys data set

```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
str(surveys)
```

Calculate the mean of the year column using pipes
```{r}
surveys$year %>% mean()
```

calculate the mean of the weight column
```{r}
surveys$weight %>% mean(na.rm = TRUE)
```

```{r}
surveys1 <- select(surveys, year, month, day, year)
str(surveys1)
```

```{r}
surveys2 <- select(surveys, year, species_id, weight)
str(surveys2)
```

```{r}
surveys2 <- mutate(surveys2, weight_kg = weight/1000)
str(surveys2)
```

```{r}
surveys2 <- filter(surveys2,!is.na(weight_kg))
str(surveys2)
```

```{r}
surveys2 <- select(surveys2, year, species_id, weight_kg)
colnames(surveys2)
surveys2[, c(1,3)]
surveys2[, c("year", "weight_kg")]
```

```{r}
surveys2_filtered <- filter(surveys2, species_id == "SH")
str(surveys2_filtered)
```

## excerise 2

1. redo all of the above but using pipes
```{r}
read.csv(file = "../data-raw/surveys (1).csv") %>% select(year, month, day, species_id) -> surveys1
str(surveys1)
surveys %>% select(year, species_id, weight) %>%
  mutate(weight_kg = weight/1000) %>%
  filter(!is.na(weight)) %>%
  select(year, species_id, weight_kg) %>%
  filter(species_id == "SH") -> surveys_final
str(surveys_final)

```

## excerise 3
```{r}
read.csv(file = "../data-raw/surveys (1).csv") %>%
  filter(species_id == "DS", !is.na(weight)) -> ds_data
str(ds_data)
ds_data %>% arrange(year) -> ds_data_by_year
str(ds_data_by_year)
ds_data_by_year %>% select(year, weight) -> ds_weight_by_year
str(ds_weight_by_year)
```