---
title: "writing functions"
author: "Davinder Singh"
date: "2023-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
## in class excerise
```{r}
get_mass_from_length_theropoda <- function(length){
  mass <- 0.73 * length ^ 3.63
  return(mass)
}
```

```{r}
get_mass_from_length_theropoda(16)
```
```{r}

```


## Use and modify function excerise
```{r}
get_mass_from_length_theropoda <- function(length = 16){
  mass <- 16 * length ^ 3.63
  return(mass)
}
```

```{r}
get_mass_from_length <- function(length){
  mass <- a * length ^ b
  return(mass)
}
```



```{r}
get_mass_from_length <- function(a, b, length){
  mass <- a * length ^ b
  return(mass)
}
get_mass_from_length(214.44, 1.46, 26)
```



## writing funcitons excerise
```{r}
 #convert_pounds_to_grams <- function(pounds = 1) {
     #grams = 453.6 * pounds
     #return(grams)
# }
#convert_pounds_to_grams() # object naming 
 # {
   #grams = 453.6 * pounds
   #return(grams)
  #} #setting fuction up
# function(x) # function varaible
```

```{r}
 convert_pounds_to_grams <- function(grams, pounds) {
     pounds <- 3.75
     grams = 453.6 * pounds
     return(grams)
 }
convert_pounds_to_grams(453.6, 3.75)
```






## why do we use functions in programming?

- we already use functions a lot. Basically everything we run in R is a function.
-We already use functions a lot. Basically everything we run in R is a function.
-Reduce the amount of code we write
-No need to repeat code, so less chance of repeating errors  or introducing new errors by miss-typing code.
- Helps with reproducibility: no need to repeat code allows others (and future self) to run code in a straightfoward way, with accuracy
-Makes it easier to write modeular code that you can reuse for other projects
- Makes it easier to run and remember it.
- It allows to organize analysis by putting code intended to run something together in the same function.


## Excerise 4
```{r}
convert_kg_to_pounds <- function(kg = 0.45){
  pounds = 2.205 * kg
  return(pounds)
}
convert_kg_to_pounds()
```

```{r}
convert_kg_to_pounds(50)
```

```{r}
get_mass_from_length(a = 10.95, b = 2.64, length = 12) %>% 
  convert_kg_to_pounds() 

```

```{r}
length_in_pounds <- function(length, a, b){
  get_mass_from_length(a = a, b = b, length = length) %>% 
  convert_kg_to_pounds() 
}
```

```{r}
length_in_pounds(33, 5.6, 6.7)
```

## excerise 5
```{r}
surveys <- read.csv(file = "../data-raw/surveys (1).csv")
species <- read.csv(file = "../data-raw/species.csv" )
plots <- read.csv(file = "../data-raw/plots.csv")
intersect(colnames(surveys), colnames(species))
```

```{r}
a = inner_join(surveys, species, by = "species_id")
head(a)
```

```{r}
my_function <- function(a, genus_name = "Dipodomys", species_name = "merriami") {
  a %>% 
  group_by(year) %>% 
  filter(genus == genus_name | species == species_name) %>% 
  summarize(count = n()) -> output
  return(output) 
}
my_function(a = a)
```



```{r}
my_function(a =a) %>% 
  ggplot(mapping = aes(x = year, y = count)) +
  geom_line(color = "blue") 
```
```{r}
my_function(a = a, "Chaetodipus", "penicillatus") %>% 
  ggplot(mapping = aes(x = year, y = count)) +
  geom_point(size = 1, color = "blue" ) +
  geom_line(size = 2, color = "blue") +
  labs(x = "year", y = "Number of Individuals")
```




