---
title: "excerises"
author: "Davinder Singh"
date: "2023-04-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```
we 'apply' functions allow us to apply a function to a vector or list of values iteratively
This helps minimize errors in code and makes the analyses more efficent

with lapply() and sapply() functions, we can only provide one argument to iterate on
'sapply()' functions simplifies the output to a vector (or the simplest data structure possible), while 'lapply()' returns an output in the form of a list 

with 'mapply()' we can provide multiple arugements to iterate on



# Calculating the mass of a bunch of dinosaurs
```{r}
mass_from_length_theropoda <- function(length){
    mass <- 0.73 * length^3.63
    return(mass)
}
```


```{r}
  theropoda_lengths <- c(17.8013631070471, 20.3764452071665, 14.0743486294308, 25.65782386974, 26.0952008049675, 20.3111541103134, 17.5663244372533, 11.2563431277577, 20.081903202614, 18.6071626441984, 18.0991894513166, 23.0659685685892, 20.5798853467837, 25.6179254233558, 24.3714331573996, 26.2847248252537, 25.4753783544473, 20.4642089867304, 16.0738256364701, 20.3494171706583, 19.854399305869, 17.7889814608919, 14.8016421998303, 19.6840911485379, 19.4685885050906, 24.4807784966691, 13.3359960054899, 21.5065994598917, 18.4640304608411, 19.5861532398676, 27.084751999756, 18.9609366301798, 22.4829168046521, 11.7325716149514, 18.3758846100456, 15.537504851634, 13.4848751773738, 7.68561192214935, 25.5963348603783, 16.588285389794)
```

```{r}
mass_from_length_theropoda(length = theropoda_lengths)
```
```{r}
theropoda_masses <- mass_from_length_theropoda(length = theropoda_lengths)

my_list <- list(theropoda_masses)
second_list <- c(my_list, list(c("Luna", "Avi", "Anitia")))
second_list[[1]]

data.frame(theropoda_masses, c("Anita", "Avi", "Luna", "Maria"))
```


```{r}
mass_from_length<- function(length, a = 0.73, b = 3.63){
  mass <- a * length^b
return(mass)
}

```

```{r}
  a_values <- c(0.759, 0.751, 0.74, 0.746, 0.759, 0.751, 0.749, 0.751, 0.738, 0.768, 0.736, 0.749, 0.746, 0.744, 0.749, 0.751, 0.744, 0.754, 0.774, 0.751, 0.763, 0.749, 0.741, 0.754, 0.746, 0.755, 0.764, 0.758, 0.76, 0.748, 0.745, 0.756, 0.739, 0.733, 0.757, 0.747, 0.741, 0.752, 0.752, 0.748)
```

```{r}
  b_values <- c(3.627, 3.633, 3.626, 3.633, 3.627, 3.629, 3.632, 3.628, 3.633, 3.627, 3.621, 3.63, 3.631, 3.632, 3.628, 3.626, 3.639, 3.626, 3.635, 3.629, 3.642, 3.632, 3.633, 3.629, 3.62, 3.619, 3.638, 3.627, 3.621, 3.628, 3.628, 3.635, 3.624, 3.621, 3.621, 3.632, 3.627, 3.624, 3.634, 3.621)
```

```{r}
mass_from_length(length = theropoda_lengths, a = a_values, b = b_values)
```

```{r}
dino_data <- data.frame(theropoda_lengths, a_values, b_values) 
dino_data %>% 
  mutate(mass = mass_from_length(theropoda_lengths)) %>% 
  head() %>% 
  return(mass)

```

# conditoning on length
```{r}
mass_from_length_max <- function(length) {
  if (length < 20){
    mass <- 0.73 * length^3.63
    return(mass)
  } else{
    mass <- NA
  }
  return(mass)
}
```

```{r}
sapply(theropoda_lengths, mass_from_length_max)
```

?mapply
#a dataset with dinosaur length
```{r}
read.csv(file = "../datasceince2023/data-raw/dinosaur_lengths (1).csv") -> dinosaur
get_mass_from_length_by_name <- function(length, dinosaur){
  if (dinosaur == "Stegosauria"){
    a = 10.95
    b = 2.64
     mass <- a * length^b
  } else if (dinosaur == "Theropoda"){
    a = 0.73
    b = 3.63
     mass <- a * length^b
  } else if (dinosaur == "Sauropoda"){
    a = 214.44
    b = 1.46
     mass <- a * length^b
  } else{
  mass = NA
  }
  return(mass)
}
mapply(FUN = get_mass_from_length_by_name, dinosaur$length, dinosaur$species )
```

```{r}
dinosaur %>%
  rowwise() %>% 
  mutate(mass = get_mass_from_length_by_name(lengths, species)) -> masses
  head(masses) 
```

```{r}
masses %>%
  filter(!is.na(mass)) %>% 
ggplot() +
  geom_histogram(mapping = aes(x = mass, color = species)) +
  facet_wrap(~species)
```
??facetwrap


